Models of VOR learning in MHC knockout mice

### 2013-02-12 Notes/VORmodel.tex
Fixed mistake in multistate analysis
Included multistate figs with intermediate training strength.

\begin{table}
 \begin{center}
  \begin{tabularn}{|l|c|c|c|c|c|c|}
    \cline{1-7}
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
    \cline{1-7}
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50  &\label{tr:cascade_short} \\
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150 &\label{tr:cascade_long} \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50  &\label{tr:multistate_weak} \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20  & \label{tr:multistate_strong} \\
    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5   &\label{tr:binary} \\
    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
                                          & -0.1 & 70 & 70 &\label{tr:pooled_plenty}\\
    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
                                          & -0.1 & 70 & 70 &\label{tr:pooled_scarce}\\
    \cline{1-7}
%   \hline
%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
%    \cline{1-7}
%    \label{tr:cascade_short}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
%    \label{tr:cascade_long}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
%    \label{tr:multistate_weak}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
%    \label{tr:multistate_strong}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
%    \label{tr:binary}&
%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
%    \label{tr:pooled_plenty}&
%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \label{tr:pooled_scarce}&
%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \cline{1-7}
  \end{tabularn}
 \end{center}
  \caption{Parameters used in simulations.} \label{tab:params}
\end{table}

@@

@@

The results of simulations of the multistate model can be seen in \autoref{fig:multistate_weak} and \autoref{fig:multistate_strong}.
However, we can get some insight into this model analytically.

@@

Let us define $\Delta f^*(\beta,M)$ to be the value at which \eqref{eq:multiKOnopre} and \eqref{eq:multiKOpre} are equal.
As we would like pre-training to slow down learning in the wild-type but speed it up in the mutant, it would seem that we require $\Delta f^*(\beta,M) < \Delta f < \Delta f^*(1,M)$ (remembering once more that $\Delta f<0$ and that the wild-type corresponds to $\beta=1$).
However, as seen in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) > \Delta f^*(1,M)$, which would make this impossible.

But, all is not lost.
This is only the initial, instantaneous rate of change.
Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f < \Delta f^*(1,M)$, so that pre-training initially slows down learning for both wild-type and mutant.
We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}) drifting to the centre.
This happens sooner for the mutant than the wild-type, due to the stronger depressing transitions.
This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the mutant.

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{multistate_betastar.eps}}\label{fig:multistate_betastar}
  \item\aligntop{\includegraphics[width=7cm]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
 \end{myenuma}
 \end{center}
  \caption{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$ and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$.}\label{fig:multistate_star}
\end{figure}

>>>>

\begin{table}
 \begin{center}
  \begin{tabularn}{|l|c|c|c|c|c|c|}
    \cline{1-7}
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
    \cline{1-7}
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 50  &\label{tr:cascade_short} \\
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 150 &\label{tr:cascade_long} \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1  & 50 & 50  &\label{tr:multistate_weak} \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.17 & 10 & 30  &\label{tr:multistate_med} \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4  & 20 & 20  &\label{tr:multistate_strong} \\
    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1  & 5  & 5   &\label{tr:binary} \\
    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
                                          & -0.1 & 70 & 70 &\label{tr:pooled_plenty}\\
    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
                                          & -0.1 & 70 & 70 &\label{tr:pooled_scarce}\\
    \cline{1-7}
%   \hline
%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
%    \cline{1-7}
%    \label{tr:cascade_short}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
%    \label{tr:cascade_long}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
%    \label{tr:multistate_weak}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
%    \label{tr:multistate_strong}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
%    \label{tr:binary}&
%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
%    \label{tr:pooled_plenty}&
%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \label{tr:pooled_scarce}&
%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \cline{1-7}
  \end{tabularn}
 \end{center}
  \caption{Parameters used in simulations.} \label{tab:params}
\end{table}

@@

\resultsfig{multistate_med}{Simulation results for the multistate model with intermediate training ($\Delta f=-0.17$)}

@@

The results of simulations of the multistate model can be seen in \autoref{fig:multistate_weak}, \autoref{fig:multistate_med} and \autoref{fig:multistate_strong}.
However, we can get some insight into this model analytically.

@@

Let us define $\Delta f^*(\beta,M)$ to be the value at which \eqref{eq:multiKOnopre} and \eqref{eq:multiKOpre} are equal.
Values of $\Delta f$ that are smaller than this (stronger training) will correspond to slower learning after pre-training.
As we would like pre-training to slow down learning in the wild-type but speed it up in the mutant, it would seem that we require $\Delta f^*(\beta,M) < \Delta f < \Delta f^*(1,M)$ (remembering once more that $\Delta f<0$ and that the wild-type corresponds to $\beta=1$).
As we see in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) < \Delta f^*(1,M)$, so this range always exists.
We can see an example of this in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}.

However, this is only the initial, instantaneous rate of change.
Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f < \Delta f^*(1,M)$, so that pre-training initially slows down learning for both wild-type and mutant.
We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}) drifting to the centre.
This happens sooner for the mutant than the wild-type, due to the stronger depressing transitions.
This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the mutant, even outside the desired range of values for $\Delta f$.
As we see in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}, even when we are in this range, the period of time for which pre-training impairs learning in the wild-type can be very small.

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{multistate_betastar.eps}}\label{fig:multistate_betastar}
  \item\aligntop{\includegraphics[width=7cm]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
 \end{myenuma}
 \end{center}
  \caption{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the mutant has impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
\end{figure}
